---
title: "Case2.Presynaptic.Rmd"
author: "Oksana Sorokina"
date: '2022-06-23'
output: html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>"
)
```

```{r setup}
library(BioNAR)
library(synaptome.db)
library(ggplot2)
library(synaptome.db)
library(pander)
library(ggrepel)
library(clusterProfiler)
library(ReactomePA)
```
## Build network for the node list extracted from SynaptomeDB

```{r net_from_gml}
# Build graph from Synaptome.db
#cid<-match('Presynaptic',getCompartments()$Name) # Let's get the ID for presynaptic compartment
#cid
#t<-getAllGenes4Compartment(cid) # Now we need to collect all the gene IDs for presynaptic  compartment
#dim(t)
#head(t)
#g<-buildFromSynaptomeByEntrez(t$HumanEntrez) # finally, build the graph using respecctive gene EntrezIDs as node IDs
#summary(g)

# Load predefined graph
g <- igraph::read.graph("~/Documents/Annet_paper/Presynaptic.gml",format="gml") #graph from gml
summary(g)

# Graph from nodes found more than once
#pres <- read.table("~/Documents/Synaptic proteome paper/db/Up_March2020/Pres_DB_April21.txt", sep = "\t", header = T, stringsAsFactors = F) 
#presC <- pres[pres$count >= 2,]
#g<-buildFromSynaptomeByEntrez(presC$HUMAN.ENTREZ.ID)
#summary(g)

```
# Annotate the nodes with node attributes
## Gene name
Adding gene names to nodes.
```{r annotate_net}

g<-annotateGeneNames(g)
summary(g)
head(V(g))
head(V(g)$GeneName)
```
## Diseases
Adding diseases associations from HDO database
```{r annotate_topOnto}
afile<-system.file("extdata", "flatfile_human_gene2HDO.csv", package = "BioNAR")
dis    <- read.table(afile,sep="\t",skip=1,header=F,strip.white=T,quote="")
g <- annotate_topOnto_ovg(g, dis)
summary(g)

```
##Shanno
Adding synaptic function annotated from Shanno et al.,paper
```{r annotate_Shanno}
sfile<-system.file("extdata", "SCH_flatfile.csv", package = "BioNAR")
shan   <- read.table(sfile,sep="\t",skip=1,header=F,strip.white=T,quote="")
g<-annotate_SCHanno(g,shan)
summary(g)
```
## Presynaptic function annotation
This is manually curated presynaptic genes specific functional annotation based on the paper of Boyken at al., 2019

```{r annotate_Chua, eval=FALSE}
sfile<-system.file("extdata", "PresynAn.csv", package = "BioNAR")
pres <- read.csv(sfile,skip=1,header=F,strip.white=T,quote="")
g <- AnNet:::annotate_CHUA(g, pres)
summary(g)
```
## GO
Adding functionality from GO: BP, MF,CC
```{r annotate_go}
sfile<-system.file("extdata", "flatfile.go.BP.csv", package = "BioNAR")
goBP <- read.table(sfile,sep="\t",skip=1,header=F,strip.white=T,quote="")
g <- annotate_go_bp(g, goBP)
summary(g)

sfile<-system.file("extdata", "flatfile.go.MF.csv", package = "BioNAR")
goMF <- read.table(sfile,sep="\t",skip=1,header=F,strip.white=T,quote="")
g <- annotate_go_mf(g, goMF)
summary(g)

sfile<-system.file("extdata", "flatfile.go.CC.csv", package = "BioNAR")
goCC <- read.table(sfile,sep="\t",skip=1,header=F,strip.white=T,quote="")
g <- annotate_go_cc(g, goCC)
summary(g)

```
## Check the netwok's degree distribution

```{r powerLaw,fig.height=9,fig.width=9}
pFit <- FitDegree( as.vector(igraph::degree(graph=g)),plot=TRUE,WIDTH=2480, HEIGHT=2480)
```
#Esimate node centrality measures
## Estimate centrality measures with values added as node attributes. 
Centrality measures are as following:DEG - degree, BET - betweenness, CC - clustering coefficient, SL - semilocal centrality, mnSP - mean shortest path, PR - page rank, sdSP - standard deviation of the shortest path

```{r graph_cent}
g <- calcCentrality(g)
summary(g)
mc <- getCentralityMatrix(g)
head(mc)
```
## Clustering
To compare different clustering algorithms summary matrix can be calculated with the maximum Modularity obtained (mod), number of detected communities (C), the number of communities with size (Cn1) equal to 1, the number of communities >= 100 (Cn100), the fraction of edges lying between communities (mu), the size of the smallest community (Min. C), the size of the largest community (Max. C), the average ( Mean C), median (Median C), first quartile (1st Qu. C), and third quartile (3rd Qu. C) of the community size. 
```{r clusterSummary}
g <- calcAllClustering(g)
m<-clusteringSummary(g,att=c('lec','wt','fc','infomap','louvain','sgG1','sgG2','sgG5', 'spectral'))
pander(m)

```
## Estimate clustering similiarity by RMI (Newman’s Reduced Mutual Information)
```{r rmi}

library(clustAnalytics)
alg<-c('lec','wt','fc','infomap','louvain','sgG1','sgG2','sgG5','spectral')

res <- matrix(0, nrow = length(alg), ncol = length(alg))
for (i in 1:length(alg)){
for (j in 1:length(alg)){
mi<-as.numeric(vertex_attr(g,alg[i]))
mj<-as.numeric(vertex_attr(g,alg[j]))
res[i,j] <- reduced_mutual_information(mi, mj, normalized = TRUE)
}}
pander(res)
colnames(res) <- alg
row.names(res) <- alg
library(corrplot)
#corrplot(res, method = "square", order = 'hclust')
library(gplots)
heatmap.2(
res, distfun = dist,
hclustfun = hclust,
col="heat.colors")
```
## Cluster enrichment for specific algorithm (louvain)

```{r enrich}
mem<- data.frame(names=V(g)$name,membership=as.numeric(V(g)$louvain))
head(mem)

Clist <- list()
for(i in 1:max(mem$membership)) {
cl <- mem[mem$membership == i,]
genes <- cl$names
Clist[[paste0("X",i)]] <- genes
}
head(Clist)

xx <- compareCluster(Clist, fun="enrichKEGG",
                     organism="hsa", pvalueCutoff=0.05)#compare cluster enichment against the KEGG database
#pdf("compareCl_louvain_KEGG.pdf", width = 12, height = 18)
dotplot(xx)
#dev.off()

goe <- compareCluster(Clist, fun="enrichGO",ont = "BP",
OrgDb    = org.Hs.eg.db, pvalueCutoff=0.01)#compare cluster enrichment against the GO BP database
#pdf("compareCl_louvain_GOBP.pdf", width = 12, height = 30)
dotplot(goe)
#dev.off()

re <- compareCluster(Clist, fun="enrichPathway", organism = "human", pvalueCutoff=0.01)
#pdf("compareCl_louvain_Reactome.pdf", width = 12, height = 20)
dotplot(re)
#dev.off()

```
## Get higher quality figure for selected clusters

```{r figure}

goe5 <- compareCluster(Clist[1:5], fun="enrichGO",ont = "BP",
OrgDb    = org.Hs.eg.db, pvalueCutoff=0.01)#compare cluster enrichment against the GO BP database

names(goe5@compareClusterResult)
grep(".*transesterification.*",goe5@compareClusterResult$Description) #identify unnecessary term
grep(".*transesterification.*",goe5@compareClusterResult$Description,value = TRUE)

goe5cl<-goe5
goe5cl@compareClusterResult<-goe5cl@compareClusterResult[-664,]

dotplot(goe5cl) -> x
#pdf("compareCl_louv5_goBP.pdf", width = 8, height = 10)
x + scale_color_gradientn(colours=c("#f7ca64", "#46bac2", "#7e62a3"), trans = "log10",guide=guide_colorbar(reverse=TRUE, order=1)) 
#dev.off()

```

## Consensus matrix
For assessing the robustness of the clustering randomization study could be performed, which applies the same clustering algorithm to N perturbed networks and returns the consensus matrix where each pair of nodes will be assigned the probability to belong to the same cluster.
```{r cons_mat}
alg = "louvain"
conmat <- makeConsensusMatrix(g, N=100,alg = alg, type = 2, mask = 10,reclust = FALSE, Cnmin = -1, Cnmax = 10)#Build consensus matrix for louvain clustering

```

##Consensus matrix value distribution
Consensus matrix value distribution could be visualised in the following way:
```{r plot.conmat.ecdf,fig.height=8,fig.width=8}
steps <- 100
Fn  <- ecdf(conmat[lower.tri(conmat)])
X<-seq(0,1,length.out=steps+1)
cdf<-Fn(X)
dt<-data.frame(cons=X,cdf=cdf)
ggplot(dt,aes(x=cons,y=cdf))+geom_line()+
      theme(            
        axis.title.x=element_text(face="bold",size=rel(2.5)),
        axis.title.y=element_text(face="bold",size=rel(2.5)),
        legend.title=element_text(face="bold",size=rel(1.5)),
        legend.text=element_text(face="bold",size=rel(1.5)),
        legend.key=element_blank())+
    theme(panel.grid.major = element_line(colour="grey40",size=0.2),
          panel.grid.minor = element_line(colour="grey40",size=0.1),
          panel.background = element_rect(fill="white"),
          panel.border = element_rect(linetype="solid",fill=NA))
```
## Cluster robustness
Cluster robustness assess the robustness of the obtained clusters and could help to evaluate the "goodness" of chosen clustering algorithm.
```{r clcons}
clrob<-getRobustness(g, alg = alg, conmat)
pander(clrob)
```

# Bridgeness 
Bridging proteins are known to interact with many neighbours simultaneously, organise function inside communities they belong to, but also affect/influence other communities in the network (Nepusz et al., 2008).Bridgeness is estimated from the consensus clustering estimated above and vertex degree to calculate the vertex’s community membership, i.e. the probability of the specific node to belong to every communities obtained by given clustering algorithm.
The Bridgeness measure lies between 0, implying a vertex clearly belongs in a single community, and 1, implying a vertex forms a 'global bridge' across every community with the same strength. 

```{r calc.bridge}
br<-getBridgeness(g,alg = alg,conmat)
pander(head(br))

```

## Bridgeness plot
Semi-local centrality measure (Chen et al., 2011) also lies between 0 and 1 indicating whether protein is important globally or locally. By plotting Bridgeness against semi-local centrality we can categorises the influence each protein found in our network has on the overall network structure:	Region 1, proteins having a 'global' rather than 'local' influence in the network (also been called bottle-neck bridges, connector or kinless hubs (0<Sl<0.5; 0.5<Br<1). Region 2, proteins having 'global' and 'local' influence (0.5<Sl<1, 0.5<Br<1). Region 3, proteins centred within the community they belong to, but also communicating with a few other specific communities (0<Sl<0.5; 0.1<Br<0.5). Region 4, proteins with 'local' impact , primarily within one or two communities (local or party hubs, 0.5<Sl<1, 0<Br<0.5).
```{r plot.bridgeness,fig.height=8,fig.width=8}
scale <- function(x, VALUE=NULL){
  
  x = as.numeric(as.vector(x))
  
  xmin <- min(x,na.rm=T)
  xmax <- max(x,na.rm=T)
  
  if( is.null(VALUE) ){
    
    x  <- x-xmin
    x  <- ifelse(!is.na(x), x/(xmax-xmin), NA) 
    
    return(x)
  }
  
  value = as.numeric(as.vector(value)[1])
  value = value-xmin
  value = ifelse(!is.na(value), value/(xmax-xmin), NA) 
  return(value)
}
#VIPs=c('8495','22999','8927','8573','26059','8497','27445','8499') #Proteins annotated with "Active zone" functionality
VIPs=c('163','3312','2775','60','112755','535','6812','478','529','6857','2782','7846','6844','4905','1759','989','6616','203068','526','6714','2885','3265','3845','999') #Proteins top found + SRC, HRAS, NRAS, GRB2
# VIPs=c('81876','10890','51552','5874','5862','11021','54734','5865','5864',
#            '9522','192683','10067','10396','9296','527','9114','537','535',
#            '528','51382','534','51606','523','80331','114569','127262','57084',
#            '57030','388662','6853','6854','8224','9900','9899','9145','9143',
#            '6855','132204','6857','127833','6861','529','526','140679','7781',
#            '81615','6844','6843')
indx   <- match(V(g)$name,VIPs)
group <- ifelse( is.na(indx), 0,1)
MainDivSize <- 0.8
xmin        <- 0
xmax        <- 1
ymin        <- 0
ymax        <- 1
Xlab <- "Semilocal Centrality (SL)" 
Ylab <- "Bridgeness (B)"
X    <- as.numeric(igraph::get.vertex.attribute(g,"SL",V(g)))
X    <- scale(X)
Y       <- as.numeric(as.vector(br[,dim(br)[2]])) 
lbls <- ifelse(!is.na(indx),V(g)$GeneName,"")
dt<-data.frame(X=X,Y=Y,vips=group,entres=V(g)$name,name=V(g)$GeneName)
dt_vips<-dt[dt$vips==1,]
dt_res<-dt[dt$vips==0,]
##--- baseColor of genes
baseColor="royalblue2"
ggplot(dt,aes(x=X,y=Y,label=name))+#geom_point()+
    geom_point(data=dt_vips,
               aes(x=X,y=Y),colour=baseColor,size = 7,shape=15,show.legend=F)+
    geom_point(data=dt_res,
               aes(x=X,y=Y, alpha=(X*Y)), size = 3,shape=16,show.legend=F)+
    geom_label_repel(aes(label=as.vector(lbls)),
                     fontface='bold',color='black',fill='white',box.padding=0.1,
                     point.padding=NA,label.padding=0.15,segment.color='black',
                     force=1,size=rel(3.8),show.legend=F,max.overlaps=200)+
      labs(x=Xlab,y=Ylab,title=sprintf("%s",alg))+
    scale_x_continuous(expand = c(0, 0), limits = c(xmin, xmax)) + 
    scale_y_continuous(expand = c(0, 0), limits = c(ymin, ymax))+
    theme(            
        axis.title.x=element_text(face="bold",size=rel(2.5)),
        axis.title.y=element_text(face="bold",size=rel(2.5)),
        legend.title=element_text(face="bold",size=rel(1.5)),
        legend.text=element_text(face="bold",size=rel(1.5)),
        legend.key=element_blank())+
    theme(panel.grid.major = element_line(colour="grey40",size=0.2),
          panel.grid.minor = element_line(colour="grey40",size=0.1),
          panel.background = element_rect(fill="white"),
          panel.border = element_rect(linetype="solid",fill=NA))+
        geom_vline(xintercept=0.5,colour="grey40",size=MainDivSize,linetype=2,show.legend=F)+
    geom_hline(yintercept=0.5,colour="grey40",size=MainDivSize,linetype=2,show.legend=F)

```
##Interactive view of bridgeness plot
```{r plotly,fig.height=10,fig.width=10}
library(plotly)
p<-ggplot(dt,aes(x=X,y=Y,label=name))+#geom_point()+
    geom_point(data=dt_vips,
               aes(x=X,y=Y),colour=baseColor,shape=15,show.legend=F)+
    geom_point(data=dt_res,
               aes(x=X,y=Y, alpha=(X*Y)),shape=16,show.legend=F)+
      labs(x=Xlab,y=Ylab,title=sprintf("%s",alg))+
    scale_x_continuous(expand = c(0, 0), limits = c(xmin, xmax)) + 
    scale_y_continuous(expand = c(0, 0), limits = c(ymin, ymax))+
    theme(            
        axis.title.x=element_text(face="bold",size=rel(2.5)),
        axis.title.y=element_text(face="bold",size=rel(2.5)),
        legend.title=element_text(face="bold",size=rel(1.5)),
        legend.text=element_text(face="bold",size=rel(1.5)),
        legend.key=element_blank())+
    theme(panel.grid.major = element_line(colour="grey40",size=0.2),
          panel.grid.minor = element_line(colour="grey40",size=0.1),
          panel.background = element_rect(fill="white"),
          panel.border = element_rect(linetype="solid",fill=NA))+
        geom_vline(xintercept=0.5,colour="grey40",size=MainDivSize,linetype=2,show.legend=F)+
    geom_hline(yintercept=0.5,colour="grey40",size=MainDivSize,linetype=2,show.legend=F)
ggplotly(p)
```
# Disease/annotation pairs
Given that Disease associated genes are connected within the graph, the common question is to check whether the networks for two different diseases are overlapping, which may indicate the common molecular mechanisms. Same is valid for any pair of annotations, e.g. one would ask if two different biological functions are related.
To address the question, we have utilised the algorithm from Menche et al, which estimates the minimal shortest paths between two distinct annotations and compare it to the randomly annotated graph.
Below example shows the estimation of disease separation for the following diseases: DOID:10652 (Alzheimer's_disease),  (bipolar_disorder), DOID:12849 (autistic_disorder), DOID:1826 (epilepsy)
Command calcDiseasePairs quickly estimates the two annotation separation on the graph and compares it with one randomly reannotated graph. This could be used for initial guess of the relationships between the annotations.
To assess the significance of the obtained separation values the command runPermDisease should be used, where the user can dfine the number of permutations. The command execution will take a while, depending on this number and return the table with p-value, p.adjusted, q-value and Bonferroni test.
```{r disPairs,warning=FALSE,message=FALSE}

p <- calcDiseasePairs(
    g,
    name = "TopOntoOVGHDO               ID",
    diseases = c("DOID:10652","DOID:3312","DOID:12849", "DOID:0060041", "DOID:9255", "DOID:12858","DOID:1059", "DOID:2377", "DOID:14330", "DOID:5419"),
    permute = "r"
)
pander(p$disease_separation)

r <- runPermDisease(
    g,
    name = "TopOntoOVGHDOID",
    diseases = c("DOID:10652","DOID:3312","DOID:12849", "DOID:0060041", "DOID:9255", "DOID:12858","DOID:1059", "DOID:2377", "DOID:14330", "DOID:5419", "DOID:1826"),
    Nperm = 10000,
    alpha = c(0.05, 0.01, 0.001)
)

pander(r$Disease_overlap_sig)
dt <- as.data.frame(r$Disease_overlap_sig)
dt$lg1val <- -log10(dt$`q-value`)

write.table(dt, file = "Dis_overlap_10000_1.txt", sep = "\t", row.names = F)
qval <- read.table("~/Documents/Annet_paper/dis_log10qval_1.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)

ggplot(qval, aes(x=X.log10.q.val., y=Pair)) + 
geom_point(col="steelblue", size=3) + geom_vline(xintercept=1.3, linetype="dashed", color = "red")  +
labs(title="Disease overlap", subtitle="Presynaptic network", y="Disease pairs", x="-Log10(q-value)")
 
#> ggplot(qval, aes(x=X.log10.q.val., y=Pair)) + 
#+     geom_point(col="steelblue", size=3) + geom_vline(xintercept=1.3, linetype="dashed", color = "red")  +
#+     labs(title="Disease overlap", subtitle="Presynaptic network", y="Disease pairs", x="-Log10(q-value)"
#+ )           
```
## Cluster overrepresentation
To identify the clusters with overrepresented function or disease we introduced the functiion which calculates the overrepesentation (enrichment for specifided annotation). Based on R package fgsea.
```{r ora,warning=FALSE,message=FALSE}}
ora <- clusterORA(g, alg, name = 'TopOnto_OVG_HDO_ID', vid = "name",alpha = 0.1, col = COLLAPSE)
```

